---
title: "Project"
author: "Osama Mufti"
date: "2024-12-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

/*

Project: Data211

According to the "Alarming Average Screen Time Statistics (2024). The average American spend 7hours and 3 minutes looking at a screen each day. In this project I will be performing T-test to compare the average screen time with my daily usage of screen time.

*/


```{r libraries}
library(tidyr)
library(ggplot2)
library(readxl)
```



```{r}
#Importing the data of 14days 

screen_time <- c(
  8 + 35/60,   # Day 1
  6 + 49/60,   # Day 2
  11 + 40/60,  # Day 3
  9 + 20/60,   # Day 4
  7 + 2/60,    # Day 5
  6 + 11/60,   # Day 6
  10 + 54/60,  # Day 7
  6 + 4/60,    # Day 8
  8 + 24/60,   # Day 9
  7 + 7/60,    # Day 10
  7 + 17/60,   # Day 11
  8 + 28/60,   # Day 12
  8 + 54/60,   # Day 13
  7 + 41/60    # Day 14
)

#Creating a data frame

days <- paste("Day", 1:14)

data <- data.frame(Day = days, ScreenTime_Hours = screen_time)

data
```


#Step:3 Performing the Calculate Descriptive Statistics

```{r}

# Calculating the mean

mean_screen_time <- mean(screen_time)


# Calculating the Median

median_screen_time <- median(screen_time)

#Standard Deviation

std_screen_time<- sd(screen_time)


# Display the results
cat("Mean Screen Time:", round(mean_screen_time, 2), "hours\n")
cat("Median Screen Time:", round(median_screen_time, 2), "hours\n")
cat("Standard Deviation:", round(std_screen_time, 2), "hours\n")

```



#Step: 3

/*

Research Question:

Is my average daily Screen time significantly different from the average American's screen time of 7.05 hours"?

Null and Alternative Hypotheses:

~Null Hypothesis $(H_0)$: The mean daily screen time is equal to 7.05 hours ($\mu =7.05$)

~Alternative Hypothesis $(H_1)$: The mean daily screen time is not equal to 7.05 hours ($\mu =7.05$ )


*/

##Step:4
/*
Perform a Two-Tailed T-Test
Since our sample size is small (n=14) a t-test is best
*/
```{r}
# Perform the t-test
t_test <- t.test(screen_time, mu = 7.05, alternative = "two.sided")

# Display the t-test results
print(t_test)

```

/*
The p-value(0.02445) is less than the significance level $(a=0.05)$

Decision: Reject the null Hypothesis

There is statistically significant evidence to concluded that your average daily screen time is different from 7.05 hours
*/



##Step: 6 Visualize Your Data

```{r}
#a: Line Graph of Daily Screen Time

data$DayIndex<-1:14

ggplot(data, aes(x = DayIndex, y = ScreenTime_Hours, group =1))+
  geom_line(color="blue")+
  geom_point(color="blue")+
  geom_hline(yintercept = 7.05, linetype= "dashed", color ="red")+
   scale_x_continuous(breaks = 1:14, labels = paste("Day", 1:14)) +
  labs(title = "Daily Screen Time Over Two Weeks", x="Day", y ="Screen Time (Hours)")+
  theme_minimal()+
  theme(axis.text.x = element_text(angle =45, hjust=1))

```

##b: Histogram of Screen Time Distribution

```{r}
# Histogram with updated `linewidth` argument
ggplot(data, aes(x = ScreenTime_Hours)) +
  geom_histogram(bins = 6, fill = "cyan", color = "black", alpha = 0.7) +
  geom_vline(xintercept = mean_screen_time, linetype = "dashed", color = "blue", linewidth = 1) +
  geom_vline(xintercept = 7.05, linetype = "dashed", color = "red", linewidth = 1) +
  labs(title = "Distribution of Daily Screen Time",
       x = "Screen Time (Hours)",
       y = "Frequency") +
  theme_minimal()

```


```

